Texture Mapping 3D Models of Indoor Environments with Noisy Camera Poses

### 2013-12-30 spie2014/spie2014.tex
thesis->paper

\author{Peter Cheng
\skiplinehalf
University of California, Berkeley\\
}

@@

The first step in automated 3D modeling is the physical scanning of
the environment's geometry. An indoor modeling system must be able to
recover its pose within an environment while simultaneously
reconstructing the 3D structure of the environment itself
\cite{chen2010indoor, hz, kua2012loopclosure, liu2010indoor}. This is
known as the simultaneous localization and mapping (SLAM) problem, and
is generally solved by taking readings from laser range scanners,
cameras, and inertial measurement units (IMUs) at multiple locations
within the environment. Mounting such devices on a platform carried by
an ambulatory human provides unique advantages over static or
vehicular-based systems on wheels in terms of agility and portability,
but can also result in larger localization error
\cite{liu2010indoor}. As a result, common methods for texture mapping
generally produce poor results. In this thesis, we present an approach
to texture mapping 3D models of indoor environments in the presence of
uncertainty and noise in camera poses. In particular, we consider data
obtained from a human-operated backpack system, detailed in Section
\ref{sec:dataAcquisition}.

@@

The overall block diagram for the proposed texture mapping procedure
is shown in Figure \ref{fig:flowchart}, where the number attached to
each box indicates the section in which the concept in the box is
explained in this thesis. First, the geometry of an environment is
split into regions, each of which will be textured independently and
in parallel. For each region, we begin by selecting a set of images
that spans the entire region's surface with high resolution
imagery. We then use recovered noisy camera poses to project these
selected images onto the surface. These projections are rotated and
translated in 2D, in order to maximally align them with the surface's
geometry, as well as to each other, allowing us to handle both errors
in geometry as well as camera poses. For surfaces and camera poses at
arbitrary locations or orientations, generally horizontal surfaces, we
propose a tile-based approach for sampling high-resolution portions of
images and compositing them into a texture. In cases where cameras
have consistently perpendicular viewing angles to the surfaces under
consideration, generally vertical surfaces, we demonstrate a superior
approach based on a shortest-path computation that leads to more
seamless textures.

The remainder of the thesis is organized as follows. Section
\ref{sec:dataAcquisition} provides background on the data acquisition
system and describes a region segmentation procedure. Section
\ref{sec:relatedWork} covers existing approaches to image stitching,
and their performance on our datasets. Section
\ref{sec:imageSelection} explains how to downsample the set of
available images by selecting those with the best orientation and
distance from surfaces under consideration. Section
\ref{sec:2dAlignment} contains the proposed approach towards 2D image
alignment, followed by Section \ref{sec:imageCompositing}, which
describes two methods of selecting and compositing images to create
the final texture. Sections \ref{sec:results} and \ref{sec:conclusion}
contain results and conclusions.

@@

This section contains background information describing the
backpack-mounted system and its operation. The hardware and operation
of the backpack system, as well as the postprocessing of acquired
data, play an important role in motivating the approaches described in
this thesis. The following two sections will focus on image and raw
data acquisition, and a method of partitioning recovered environment
geometry, respectively.

@@

A CAD model of the backpack-mounted system is shown in Figure
\ref{fig:backpack}. This backpack system has a number of laser range
scanners as well as 2 cameras facing left and right, each equipped
with fisheye lenses reaching an approximately 180$^{\circ}$ field of
view each and taking photos at a rate of 5 Hz \cite{liu2010indoor}. In
order to collect data, an operator wears the backpack system and
traverses the environment to be scanned, walking in straight lines at
a standard pace and making an effort to walk parallel to all walls in
the area. Multiple localization and loop-closure algorithms are then
applied to the raw data collected by the onboard sensors
\cite{chen2010indoor, kua2012loopclosure, liu2010indoor}, and the
backpack is localized\footnote{In this thesis, we use the terms
  localization and pose recovery interchangeably, in that they both
  refer to recovering position and orientation.}  over its data
collection period. This involves recovering the 6 degrees of freedom
for the backpack itself as well as the cameras rigidly mounted on it.
These results are noisy, which motivates the approaches in this
thesis. Once this is complete, the data from the laser range scanners
is used to generate a 3D point cloud of the surrounding
environment. This point cloud is used to construct either a
low-resolution planar model, such as for usage in basic navigation or
energy analysis \cite{sanchez2012point, turnerfloorplan}, or a
high-resolution, detailed mesh, for applications such as accurate
remodeling or virtual walkthroughs \cite{turnerwatertight}.
Regardless of resolution, size, or detail, the model, consisting of
geometric surfaces in 3D space, along with the set of images captured
by the backpack's cameras and their noisy 3D poses, are the input to
the texture mapping problem.

@@

Similar to the image position refinement procedure in Section
\ref{sec:robustSIFTFeatureMatching}, exposure compensation is
performed simulatenously across all images present in a single
region. As before, this is solved as a least-squares optimization
problem with pair-wise observations. In this case, observations do not
need to be weighted, and so the formulation is
$\textrm{min}_{\vec{\beta}} ||(A \vec{\beta} - \vec{\gamma})||_2^2
$. An example setup for 3 images, each with 2 overlapping pixels, is
shown below.
\[
A =
\begin{pmatrix}
  P_{11} & -P_{12} & 0\\
  P_{21} & -P_{12} & 0\\
  0 & P_{32} & -P_{33}\\
  0 & P_{42} & -P_{43}\\
  1 & 1 & 1\\

\end{pmatrix}\quad
\vec{\beta} =
\begin{pmatrix}
  G_1, \\ G_2, \\ G_3
\end{pmatrix}
\vec{\gamma} =
\begin{pmatrix}
  0, \\ 0, \\ 0, \\ 0, \\ 3
\end{pmatrix}
\]

In this problem, $\beta$ is the vector of gains $G_i$ to be solved
for.  The observations, represented by $A$ and $\gamma$, correspond to
equalizing the brightness for each pixel present in two images. For
instance, for pixel $P_x$ which is present in two overlapping images
$I_i$ and $I_j$, the goal is to find gains $G_i$ and $G_j$ such that
$G_iP_{xi} - G_jP_{xj} = 0$, where $P_{xy}$ is the value of pixel
$P_x$ in image $I_y$. Thus, $\gamma$ contains all zeros, while each
row of $A$ contains a $P_{xi}$ value in the $ith$ column and a
$-P_{xj}$ value in the $jth$ column, and zeros elsewhere. This problem
is not constrained, as it allows all computed gains to be scaled by
any amount. To keep gains at reasonable values, a final observation is
added: $\Sigma G_i = N$, where $N$ is the number of images. This has
the effect of setting the average of all gains to 1, and is
represented by adding a row of ones to $A$, and a corresponding $N$
value to $\gamma$.

The result of computing and applying gains for each image can be seen
for the same ceiling location in Figure \ref{fig:exposureDiff}(b). As
compared to Figure \ref{fig:exposureDiff}(a), the effect is clear, as
dark/bright patches are no longer present, and the overall brightness
of the image has not changed significantly.

@@

\subsection{Examples}
\label{sec:examples}
As mentioned in Section \ref{sec:dataAcquisition}, the texture mapping
procedure in this thesis can be applied both to low-resolution and
high-resolution models. Low resolution models have an advantage in
their simplicity, and in that only important environmental features
are represented. As a result, images are generated for large sections
of buildings, and texture continuity is maintained over large
areas. High resolution models have an obvious advantage in terms of
the amount of detail that can be reconstructed, but with poor region
partitioning, as described in Section \ref{sec:geometryPartitioning},
or with high camera pose or geometry reconstruction error, 3D elements
in the environment geometry may not match up well with their imaged
counterparts.

@@

\subsection{Runtime}
As mentioned earlier, our approach is quite efficient. The top wall in
Figure \ref{fig:indivPlanes}(a) was generated with 7543 $\times$ 776
pixels, and spans a 40-meter long wall. Given 41000 input images in
the entire dataset, a 2.8GHz dual-core consumer-grade laptop takes
under a second to choose 36 candidate images, followed by a minute to
perform image projections, image alignment, and the shortest path
texturing method. The tile-caching approach takes roughly comparable
time, and over 75\% of the time for either method is spent on
calculating image projections, which are cached over multiple runs,
and also could be performed as a preprocessing step instead. While not
real-time, the process is capable of generating fast updates after
changes in various parameters or modifications to input data, and if
integrated directly into a 3D modeling system, could provide quick
visual feedback as data is collected.

\subsection{Visualization}
Our full models consist of an input model file, generated textures,
and a mapping of image points to 3D model vertices. The textured
models shown throughout this thesis range from 20 MB in size to over
400 MB in a compressed format, with textures of sizes over 500
megapixels. These models are visualized using the OpenSceneGraph
toolkit \cite{openscenegraph}, which allows for export to many common
model formats, as well as interactive visualization, even in web
browsers or mobile devices.

@@

In this thesis, we have developed an approach to texture mapping
models of indoor environments with noisy camera localization data. We
are able to refine image locations based on geometry references and
feature matching, and robustly handle outliers. Using the tile-based
mapping approach, we can texture both large planar features as well as
smaller, more complex surfaces. We also implemented a shortest path
texturing method that produces seamless textures on planes where
multiple head-on images are available. Both of these approaches are
highly modular, and easily tunable for similar systems across multiple
environments.

@@

\begin{figure}
  \centering{
    \includegraphics[width=6.2in,height=3in]{houston2d1.png} \\
    \includegraphics[width=6.2in,height=3in]{houston2d2.png} \\
    \includegraphics[width=6.2in,height=3in]{houston3.png}
  }
  \caption{A textured low-resolution model from the floor plan extrusion approach.}
  \label{fig:e2d1}
\end{figure}

>>>>

\author{Peter Cheng\supit{a}, Michael Anderson\supit{a}, Stewart He\supit{b}, and Avideh Zakhor\supit{a}
\skiplinehalf
\supit{a}University of California, Berkeley; \\
\supit{b}University of California, Davis
}

@@

The first step in automated 3D modeling is the physical scanning of
the environment's geometry. An indoor modeling system must be able to
recover its pose within an environment while simultaneously
reconstructing the 3D structure of the environment itself
\cite{chen2010indoor, hz, kua2012loopclosure, liu2010indoor}. This is
known as the simultaneous localization and mapping (SLAM) problem, and
is generally solved by taking readings from laser range scanners,
cameras, and inertial measurement units (IMUs) at multiple locations
within the environment. Mounting such devices on a platform carried by
an ambulatory human provides unique advantages over static or
vehicular-based systems on wheels in terms of agility and portability,
but can also result in larger localization error
\cite{liu2010indoor}. As a result, common methods for texture mapping
generally produce poor results. In this paper, we present an approach
to texture mapping 3D models of indoor environments in the presence of
uncertainty and noise in camera poses. In particular, we consider data
obtained from a human-operated backpack system, detailed in Section
\ref{sec:dataAcquisition}.

@@

The overall block diagram for the proposed texture mapping procedure
is shown in Figure \ref{fig:flowchart}, where the number attached to
each box indicates the section in which the concept in the box is
explained in this paper. First, the geometry of an environment is
split into regions, each of which will be textured independently and
in parallel. For each region, we begin by selecting a set of images
that spans the entire region's surface with high resolution
imagery. We then use recovered noisy camera poses to project these
selected images onto the surface. These projections are rotated and
translated in 2D, in order to maximally align them with the surface's
geometry, as well as to each other, allowing us to handle both errors
in geometry as well as camera poses. For surfaces and camera poses at
arbitrary locations or orientations, generally horizontal surfaces, we
propose a tile-based approach for sampling high-resolution portions of
images and compositing them into a texture. In cases where cameras
have consistently perpendicular viewing angles to the surfaces under
consideration, generally vertical surfaces, we demonstrate a superior
approach based on a shortest-path computation that leads to more
seamless textures.

The remainder of the paper is organized as follows. Section
\ref{sec:dataAcquisition} provides background on the data acquisition
system and describes a region segmentation procedure. Section
\ref{sec:relatedWork} covers existing approaches to image stitching,
and their performance on our datasets. Section
\ref{sec:imageSelection} explains how to downsample the set of
available images by selecting those with the best orientation and
distance from surfaces under consideration. Section
\ref{sec:2dAlignment} contains the proposed approach towards 2D image
alignment, followed by Section \ref{sec:imageCompositing}, which
describes two methods of selecting and compositing images to create
the final texture. Sections \ref{sec:results} and \ref{sec:conclusion}
contain results and conclusions.

@@

This section contains background information describing the
backpack-mounted system and its operation. The hardware and operation
of the backpack system, as well as the postprocessing of acquired
data, play an important role in motivating the approaches described in
this paper. The following two sections will focus on image and raw
data acquisition, and a method of partitioning recovered environment
geometry, respectively.

@@

A CAD model of the backpack-mounted system is shown in Figure
\ref{fig:backpack}. This backpack system has a number of laser range
scanners as well as 2 cameras facing left and right, each equipped
with fisheye lenses reaching an approximately 180$^{\circ}$ field of
view each and taking photos at a rate of 5 Hz \cite{liu2010indoor}. In
order to collect data, an operator wears the backpack system and
traverses the environment to be scanned, walking in straight lines at
a standard pace and making an effort to walk parallel to all walls in
the area. Multiple localization and loop-closure algorithms are then
applied to the raw data collected by the onboard sensors
\cite{chen2010indoor, kua2012loopclosure, liu2010indoor}, and the
backpack is localized\footnote{In this paper, we use the terms
  localization and pose recovery interchangeably, in that they both
  refer to recovering position and orientation.}  over its data
collection period. This involves recovering the 6 degrees of freedom
for the backpack itself as well as the cameras rigidly mounted on it.
These results are noisy, which motivates the approaches in this
paper. Once this is complete, the data from the laser range scanners
is used to generate a 3D point cloud of the surrounding
environment. This point cloud is used to construct either a
low-resolution planar model, such as for usage in basic navigation or
energy analysis \cite{sanchez2012point, turnerfloorplan}, or a
high-resolution, detailed mesh, for applications such as accurate
remodeling or virtual walkthroughs \cite{turnerwatertight}.
Regardless of resolution, size, or detail, the model, consisting of
geometric surfaces in 3D space, along with the set of images captured
by the backpack's cameras and their noisy 3D poses, are the input to
the texture mapping problem.

@@

Similar to the image position refinement procedure in Section
\ref{sec:robustSIFTFeatureMatching}, exposure compensation is
performed simulatenously across all images present in a single
region. To begin, a relative gain $R_{ij}$ is computed for each pair
of overlapping images $I_i$ and $I_j$, which have a set of pixels
common to both images, labeled as $P_{ij}$. The relative gain $R_{ij}$
is then obtained by calculating the scaling factor between the average
intensity of $P_{ij}$ in $I_i$ and the average intensity of $P_{ij}$
in $I_j$.

As before, these pair-wise observations can be combined and solved as
a least-squares optimization problem. In this case, observations do
not need to be weighted, and so the formulation is
$\textrm{min}_{\vec{b}} ||(X \vec{b} - \vec{a})||_2^2 $. An example
setup for 3 images, where each image overlaps with both of the other
images, is shown below.
\[
X =
\begin{pmatrix}
  R_{12} & -1 & 0\\
  R_{13} & 0 & -1\\
  0 & R_{23} & -1\\
  1 & 1 & 1\\

\end{pmatrix}\quad
\vec{b} =
\begin{pmatrix}
  G_1, \\ G_2, \\ G_3
\end{pmatrix}
\vec{a} =
\begin{pmatrix}
  0, \\ 0, \\ 0, \\ 3
\end{pmatrix}
\]

In this problem, $\vec{b}$ is the $N\times1$ vector of gains $G_i$ to
be solved for, where $N$ is the number of images.  Matrix $X$ is an
observation matrix corresponding to the relative intensities of
overlapping images. In particular, the number of rows of $X$ is one
plus the number of unique pairs of overlapping images, and the number
of columns of $X$ is equal to $N$. Each row of $X$, except the last
one, contains an $R_{ij}$ value which is placed in column $i$, as well
as a $-1$ value in column $j$, with zeroes elsewhere. As explained
above, $R_{ij}$ is the relative gain between images $i$ and $j$, and
therefore $G_iR_{ij} - G_j = 0$.  Thus, all values of $\vec{a}$, which
is a vector with size equal to the number of rows of $X$, are set to
$0$, except for the last one.

The last row of $X$ and corresponding last entry of $\vec{a}$ are used
to constrain the problem, as without them, all computed gains would be
allowed to scale by any amount. To keep gains at reasonable values,
the final observation is used: $\Sigma G_i = N$, where $N$ is the
number of images. This has the effect of setting the average of all
gains to 1, and is represented by the final row of ones in $X$, and
the corresponding $N$ value in $\vec{a}$.

For the 81 source images that compose the ceiling location in Figure
\ref{fig:exposureDiff}, the corresponding $X$ matrix has 1343 rows,
and the optimization problem takes under a second to solve, on a 2.8
GHz dual core machine.  The result of computing and applying gains can
be seen in Figure \ref{fig:exposureDiff}(b). As compared to Figure
\ref{fig:exposureDiff}(a), the effect is clear, as dark/bright patches
are no longer present, and the overall brightness of the image has not
changed significantly.

@@

\subsection{Examples}
\label{sec:examples}
As mentioned in Section \ref{sec:dataAcquisition}, the texture mapping
procedure in this paper can be applied both to low-resolution and
high-resolution models. Low resolution models have an advantage in
their simplicity, and in that only important environmental features
are represented. As a result, images are generated for large sections
of buildings, and texture continuity is maintained over large
areas. High resolution models have an obvious advantage in terms of
the amount of detail that can be reconstructed, but with poor region
partitioning, as described in Section \ref{sec:geometryPartitioning},
or with high camera pose or geometry reconstruction error, 3D elements
in the environment geometry may not match up well with their imaged
counterparts.

@@

\subsection{Runtime}
As mentioned earlier, our approach is quite efficient. The top wall in
Figure \ref{fig:indivPlanes}(a) was generated with 7543 $\times$ 776
pixels, and spans a 40-meter long wall. Given 41000 input images in
the entire dataset, a 2.8 GHz dual-core consumer-grade laptop takes
under a second to choose 36 candidate images, followed by a minute to
perform image projections, image alignment, and the shortest path
texturing method. The tile-caching approach takes roughly comparable
time, and over 75\% of the time for either method is spent on
calculating image projections, which are cached over multiple runs,
and also could be performed as a preprocessing step instead. While not
real-time, the process is capable of generating fast updates after
changes in various parameters or modifications to input data, and if
integrated directly into a 3D modeling system, could provide quick
visual feedback as data is collected.

\subsection{Visualization}
Our full models consist of an input model file, generated textures,
and a mapping of image points to 3D model vertices. The textured
models shown throughout this paper range from 20 MB in size to over
400 MB in a compressed format, with textures of sizes over 500
megapixels. These models are visualized using the OpenSceneGraph
toolkit \cite{openscenegraph}, which allows for export to many common
model formats, as well as interactive visualization, even in web
browsers or mobile devices.

@@

In this paper, we have developed an approach to texture mapping
models of indoor environments with noisy camera localization data. We
are able to refine image locations based on geometry references and
feature matching, and robustly handle outliers. Using the tile-based
mapping approach, we can texture both large planar features as well as
smaller, more complex surfaces. We also implemented a shortest path
texturing method that produces seamless textures on planes where
multiple head-on images are available. Both of these approaches are
highly modular, and easily tunable for similar systems across multiple
environments.

@@

\begin{figure}
  \centering{
    \includegraphics[width=6.2in,height=3in]{houston2d1.png} \\
    \includegraphics[width=6.2in,height=3in]{houston2d2_2.png} \\
    \includegraphics[width=6.2in,height=3in]{houston3_2.png}
  }
  \caption{A textured low-resolution model from the floor plan extrusion approach.}
  \label{fig:e2d1}
\end{figure}
